\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Journal: Ecosphere. Manuscript ID: ECS23-0342. Manuscript type: Article}}

### 2024-02-12 manuscript/manuscriptRW.Rmd
comment9: add to figure captions all abreviations from figures

**Figure 2.** Mean regeneration levels across all samples per site, plotted for the 200 sites and for each model. The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data. There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm. The boxplot shows the median (midline), interquartile range (box limits), largest and smallest value within 1.5 times interquartile range (whiskers), larger values are represented as outliers.

**Figure 3.** Shannon index of tree regeneration, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively. The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data. Boxplot features are the same as in Figure 2.

**Figure 4.** Mean Shannon index across all samples per site for observed and simulated data. Each plot shows one pattern represented by one exemplary model of each category (overpredicted, intermediate and underpredicted). The full data with the grouping of the models are shown in Figures S3 and S4. n indicates the number of models falling in each group. A) Examples for the three trends across models for species diversity at the stand level. B) Examples for the three trends across models for regeneration at the 7 cm threshold (165 sites).

**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm thresholds. Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.Boxplot features are the same as in Figure 2.

@@

**Figure 7.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of total basal area. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.

**Figure 8.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of climatic water balance. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.

**Figure 9.** Mean recruitment across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of seasonal degree-day sum. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data.

**Figure 10.** Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold. The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient. The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species. The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether. The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.

>>>>

**Figure 2.** Mean regeneration levels ($\overline{R}$) across all samples per
site, plotted for the 200 sites and for each model. The red dashed lines
shows the 25th and 75th percentiles for the 7 cm diameter threshold in
the observed data. There are two boxplots for each model where the left
lighter boxplot corresponds to 7 cm and the right darker boxplot to 10
cm. The boxplot shows the median (midline), interquartile range (box
limits), largest and smallest value within 1.5 times interquartile range
(whiskers), larger values are represented as outliers.

**Figure 3.** Mean Shannon index ($\overline{H}$) of tree regeneration, calculated by basal
area, as the mean value across the 200 samples per site for the diameter
thresholds of 7 and 10 cm, respectively. The red dashed lines shows the
25th and 75th percentiles for the 7 cm diameter threshold in the
observed data. Boxplot features are the same as in Figure 2.

**Figure 4.** Mean Shannon index ($\overline{H}$) across all samples per site for
observed and simulated data. Each plot shows one pattern represented by
one exemplary model of each category (overpredicted, intermediate and
underpredicted). The full data with the grouping of the models are shown
in Figures S3 and S4. n indicates the number of models falling in each
group. A) Examples for the three trends across models for species
diversity at the stand level. B) Examples for the three trends across
models for regeneration at the 7 cm threshold (165 sites).

**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm
thresholds. Dashed blue lines mark a ratio equal to 1, indicating no
decrease in tree regeneration between 7 and 10 cm, and a ratio equal to
1.77, corresponding to the Reineke self-thinning ratio under even-aged
conditions. Boxplot features are the same as in Figure 2.

@@

**Figure 7.** Mean recruitment ($\overline{R}$) across the 200 samples per site with the
y-axis scaled differently by model, for the 200 sites against gradients
of total basal area. The values were split into ten bins; the red lines
represent a GAM model showing the trend in the observed data.

**Figure 8.** Mean recruitment ($\overline{R}$) across the 200 samples per site with the
y-axis scaled differently by model, for the 200 sites against gradients
of climatic water balance. The values were split into ten bins; the red
lines represent a GAM model showing the trend in the observed data.

**Figure 9.** Mean recruitment ($\overline{R}$) across the 200 samples per site with the
y-axis scaled differently by model, for the 200 sites against gradients
of seasonal degree-day sum. The values were split into ten bins; the red
lines represent a GAM model showing the trend in the observed data.

**Figure 10.** Mean recruitment levels ($\overline{R}$) of the five main species in terms of
their basal area share in the observations (top row) and the performance
of each model across the environmental gradients (other rows) for the 7
cm regeneration threshold. The values shown are the mean of the 200
samples per site and across the sites in each bin (tile), with ten bins
per gradient. The sizes of the circles represent the ratio between the
regeneration basal area of the species and the total regeneration basal
area for all species. The absence of a circle indicates a zero basal
area share in the regeneration, or the absence of regeneration
altogether. The color gradient (for the models only) shows the
difference between the simulated and observed ratio of regeneration
basal area of the species and the total regeneration basal area for all
the species.
